---
title: "WBPCB Data"
author: "Rahul Konar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read #WBPCB_OFFICE_SaltLake_AQMS_Data
```{r}
my_dat<-read.csv("D:\\ISI\\Semester 6 B.Stat\\DXAQMS-1_WBPCB Office, Saltlake.csv",header=TRUE)
```

#Histogram of PM10 conc.
```{r}
hist(my_dat$pm10)
```


#Fit a regression on temp and humidity

```{r}
model<-lm(pm10~pm25,data=my_dat)
summary(model)
```


```{r}
model<-lm(pm10~pm25+ext_temp+ext_humi+so2+no2,data=my_dat)
summary(model)
```


```{r}
model<-lm(pm10~so2+no2+ext_humi+ext_temp, data = my_dat)
summary(model)
```

```{r}
model<-lm(pm10~pm25+so2, data = my_dat)
summary(model)
```

```{r}
model<-lm(pm10~ext_temp, data = my_dat)
summary(model)
```

```{r}
model<-lm(pm10~ext_humi, data = my_dat)
summary(model)
```
```{r}
model<-lm(pm10~so2+pm25,data=my_dat)
summary(model)

```

```{r}
my_dat$time <- as.POSIXct(
  paste(my_dat$date, my_dat$hour),
  format = "%Y-%m-%d %H",
  tz = "Asia/Kolkata"
)
plot(my_dat$time, my_dat$pm25,
     type = "l",
     xlab = "Time",
     ylab = "PM2.5",
     main = "Hourly PM2.5 Time Series")
```

```{r}
library(tseries)
library(lubridate)
library(dplyr)
library(urca)
my_dat<-my_dat %>%
  mutate(
    date = ymd(my_dat$date),
    datetime = date+hours(my_dat$hour)
  )%>%
  arrange(datetime)
df_pm25<-my_dat %>%
  dplyr::select(ext_temp,pm25,pm10,so2,no2,ext_humi) %>%
  na.omit()
adf_temp<-ur.df(df_pm25$ext_temp,type="trend",lags=27) #27
summary(adf_temp)
adf_pm25<-ur.df(df_pm25$pm25,type="trend",lags=24)  #24
summary(adf_pm25)
adf_pm10<-ur.df(df_pm25$pm10, type = "trend" , lags = 31) #31
summary(adf_pm10)
adf_so2<-ur.df(df_pm25$so2, type = "trend", lags = 26)
summary(adf_so2)
adf_no2<-ur.df(df_pm25$no2, type= "trend", lags =28) #28
summary(adf_no2)
adf_ext_humi<-ur.df(df_pm25$ext_humi, type = "trend", lags=26) #26
summary(adf_ext_humi)
```


For "lag+trend" model: \[\delta y_t = a_0 + \gamma y_{t-1} + a_2t + e_t\]  (Enders p.208)\\
The hypotheses:\\
$\tau :\gamma=0$\\
$\phi_3: \gamma = a_2 = 0$\\
$\phi_2: a_0 = \gamma = a_2 = 0$\\
We test all hypotheses at $5\%$ level.

```{r}
res<-residuals(adf_temp@testreg)
acf(res)
Box.test(res,lag=27,type="Ljung-Box")
```



The resulting ACF plot of residuals shows that AIC is not sufficient for choosing the lag order. ADF test validity depends on residuals being roughly uncorrelated.

```{r}
res2<-residuals(adf_pm25@testreg)
acf(res2)
```

```{r}
res3<-residuals(adf_pm10@testreg)
acf(res3)
```

```{r}
res4<-residuals(adf_so2@testreg)
acf(res4)
```

```{r}
res5<-residuals(adf_no2@testreg)
acf(res5)
```
```{r}
res6<-residuals(adf_ext_humi@testreg)
acf(res6)
```
```{r}
diff_so2<-diff(my_dat$so2)
adf_diff_so2<-ur.df(diff_so2,type="trend",lags=25)
summary(adf_diff_so2)
res7<-residuals(adf_diff_so2@testreg)
acf(res7)
```





Apart from $SO_2$ concentration time series, all other time series are stationary (i.e. they have a mean reverting behaviour, random shocks do not persist till infinity). First order differences of $SO_2$ time series is stationary.

Based on results from fitting linear trend with intercept, it seems that the linear trend may not be significant. So we fit the drift model (fitting intercept only, without trend).
```{r}
adf_temp2<-ur.df(df_pm25$ext_temp,type="drift",lags=27)
summary(adf_temp)
adf_pm25_2<-ur.df(df_pm25$pm25,type="drift",lags=24)
summary(adf_pm25)
```

We observe that the conclusions are similar when we fit intercept only for both time series. The temperature time series only rejects the null hypothesis of non-stationarity. There is no evidence of linear trend or intercept. On the other hand, the PM2.5 time series shows evidence against non-stationarity, and has significant intercept. However, the trend effect is negligible.

#VAR Model
We next fit the VAR model with lag order selected by AIC.

```{r}
library(dplyr)
library(vars)
#conflicts()
pair_pm25<-my_dat %>%
  dplyr::select(ext_temp,pm25) %>%
  na.omit()

lag<-VARselect(pair_pm25,lag.max=48,type="const")
lag$selection

p<-lag$selection["AIC(n)"]
p
var_fit<-VAR(pair_pm25, p = p, type ="const")
summary(var_fit)

  
```

# Granger Causality
```{r}
g_temp_pm25<-causality(var_fit, cause="ext_temp")
g_temp_pm25$Granger
```

```{r}
g_pm25_temp<-causality(var_fit,cause="pm25")
g_pm25_temp$Granger
```

We find that neither PM2.5 has any Granger causal effect on external temperature, and neither does external temperature have any Granger causal effect on PM2.5, for our current model.

Next, we account for the other factors and run mutivariate VAR.

```{r}
my_dat <- my_dat%>%
  dplyr::mutate(
    diff_so2 = c(NA,diff(my_dat$so2))
  )
multi_data<-my_dat %>%
  dplyr::select(ext_temp,pm25,pm10,no2,ext_humi,diff_so2) %>%
  na.omit()
lag_sel<-VARselect(multi_data, lag.max=48,type="const")
lag_sel$selection
p<-lag_sel$selection["AIC(n)"]
var_fit_2<-VAR(multi_data, p=p, type = "const")
summary(var_fit)

```

```{r}
causality(var_fit_2, cause = "ext_temp")$Granger #Does temp affect PM2.5 after controlling for everything else
causality(var_fit_2, cause = "pm25")$Granger #Does PM2.5 Granger cause temp after controlling for everything else
```

#Attempting to make causal relations between other variables.

```{r}
pair_ext_humi<-my_dat %>%
  dplyr::select(ext_temp,ext_humi) %>%
  na.omit()

lag<-VARselect(pair_ext_humi,lag.max=48,type="const")
lag$selection

p<-lag$selection["AIC(n)"]
p
var_fit2<-VAR(pair_ext_humi, p = p, type ="const")
summary(var_fit2)

```

```{r}
g_temp_humi<-causality(var_fit2, cause="ext_temp")
g_temp_humi$Granger
g_humi_temp<-causality(var_fit2,cause="ext_humi")
g_humi_temp$Granger
```

```{r}
library(dplyr)
library(vars)

# ── 1. Define variables ──────────────────────────────────────────────────────
fields <- c("ext_temp", "ext_humi", "so2", "no2", "pm10", "pm25")
pairs  <- combn(fields, 2, simplify = FALSE)   # all 15 unique pairs

# ── 2. Function: fit VAR + test Granger causality both ways ─────────────────
granger_pair <- function(var1, var2, data, lag.max = 48, ic = "AIC(n)") {
  
  df <- data %>%
    dplyr::select(all_of(c(var1, var2))) %>%
    na.omit()
  
  # Select optimal lag
  p <- VARselect(df, lag.max = lag.max, type = "const")$selection[[ic]]
  
  # Fit VAR
  fit <- VAR(df, p = p, type = "const")
  
  # Granger causality both directions
  gc_1_causes_2 <- causality(fit, cause = var1)$Granger  # var1 → var2
  gc_2_causes_1 <- causality(fit, cause = var2)$Granger  # var2 → var1
  
  tibble(
    cause    = c(var1,  var2),
    effect   = c(var2,  var1),
    opt_lag  = p,
    F_stat   = c(gc_1_causes_2$statistic, gc_2_causes_1$statistic),
    df1      = c(gc_1_causes_2$parameter[1], gc_2_causes_1$parameter[1]),
    df2      = c(gc_1_causes_2$parameter[2], gc_2_causes_1$parameter[2]),
    p_value  = c(gc_1_causes_2$p.value,    gc_2_causes_1$p.value),
    sig      = symnum(
                 c(gc_1_causes_2$p.value, gc_2_causes_1$p.value),
                 cutpoints = c(0, .001, .01, .05, .1, 1),
                 symbols   = c("***", "**", "*", ".", " ")
               ) %>% as.character()
  )
}

# ── 3. Run over all pairs ────────────────────────────────────────────────────
granger_results <- purrr::map_dfr(pairs, ~ granger_pair(.x[1], .x[2], my_dat))

# ── 4. View results ──────────────────────────────────────────────────────────
print(granger_results, n = Inf)

# Optional: filter only significant results (p < 0.05)
granger_results %>% filter(p_value < 0.05)
```

```{r}
readr::write_csv(granger_results, "D:/ISI/Semester 6 B.Stat/Statistics Comprehensive/AQMS_Data/granger_results.csv")
```